---
title: "vote_history_analysis"
output:
  pdf_document: default
  html_document: default
date: "2025-11-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, warning = FALSE, message = FALSE}
library(dplyr)
library(ggplot2)
#library(cwi) 
if (require(sf)) {
  sf::sf_use_s2(FALSE)
}
library(RSQLite)
library(DBI)
library(leaflet)
library(leaflet.extras2)
library(htmlwidgets)
library(tidygeocoder)
library(tidyverse)
library(data.table)
library(purrr)
library(leafgl)
library(lubridate)
library(reactable)
library(flextable)
library(magrittr)
library(gt)
library(htmltools)
```

```{r}
# Grab the data from sqllite connections
# setwd("C:/Users/Ruairi/Documents/Coding Projects 2025/ct04_testing")
# conn <- dbConnect(RSQLite::SQLite(), "C:/Users/Ruairi/Downloads/voters.db")
# df_voters <- dbGetQuery(conn, "SELECT * FROM voters")
# df_vh <- dbGetQuery(conn, "select * from voter_history")

df_voters_with_history <- read.csv('clean_vf_with_vh.csv')
```

### What are the best town to target?

```{r town_aggregation}
town_agg <- df_voters_with_history %>% 
  group_by(town_name) %>%
  mutate(phone_avail = ifelse(phone != '',1,0),
         active_dem_phone_avail = ifelse(phone != '' & active_dem == 1,1,0),
         likely_signers_phone_avail = ifelse(phone != '' & likely_signer == 1,1,0)) %>% 
  summarise(vf_records = n(),
            active_dems = sum(active_dem),
            likely_signers = sum(likely_signer),
            phones = sum(phone_avail),
            ad_phones = sum(active_dem_phone_avail),
            ls_phones = sum(likely_signers_phone_avail)) %>%
  mutate(ratio_likely_signers_to_act_dems = round(likely_signers/active_dems, 2),
         p_voters_active_dems = round(100*active_dems/vf_records,0),
         p_voters_likely_signers = round(100*likely_signers/vf_records,0),
         p_voters_phones = round(phones/vf_records,2),
         p_ad_phones = round(ad_phones/active_dems,2),
         p_ls_phones = round(ls_phones/active_dems, 2)) %>%
  #get rid of incorrect spelling, neighborhoods, etc 
  filter(vf_records > 400) %>%
  mutate(Tier = ifelse(likely_signers > 6000, 1, ifelse(likely_signers > 1500, 2, 3)))
```

# HTML Table of Aggregated Town Counts

```{r}
# Making a clean table
tab_town <- town_agg %>% 
  arrange(desc(likely_signers)) %>% 
  rename(Town = town_name, Total_Records = vf_records) %>%
  mutate(Total_Records = format(Total_Records, big.mark = ",", scientific = FALSE),
         Active_Dems = paste0(format(active_dems, big.mark = ",", scientific = FALSE), ' ', '(', p_voters_active_dems ,'%)'),
         Likely_Signers = paste0(format(likely_signers, big.mark = ",", scientific = FALSE), ' ', '(', p_voters_likely_signers ,'%)'), 
         Ratio_LS_to_AD = ratio_likely_signers_to_act_dems
         ) %>% select(Town, Total_Records, Active_Dems, Likely_Signers, Tier, Ratio_LS_to_AD)

tab_town_styled <- reactable(tab_town,
          striped = TRUE,
          defaultPageSize = 25,
          rowStyle = function(index) {
            # Get the value from the 'cyl' column for the current row
            x <- tab_town[index, "Tier"]
            
            # Apply a background color based on the condition
            if (x == 1) {
              list(background = "#ccffcc") # Light green for 4 cyl
            } else if (x == 2) {
              list(background = "#cce5ff") # Light blue for 6 cyl
            } else if (x == 3) {
              list(background = "#ffcccc") # Light red for 8 cyl
            }
          }
        )

saveWidget(tab_town_styled, "dems_by_town.html", selfcontained = TRUE)

town_agg %>% 
  group_by(Tier) %>% 
  summarise(Active_Dems = format(round(sum(active_dems)/1000, 0)*1000, big.mark=','), 
            High_Propensity = format(round(sum(likely_signers)/1000,0)*1000, big.mark = ","))

left_join(df_voters_with_history, town_agg %>% select(town_name, Tier), by='town_name') %>% 
  mutate(Active_Dem_w_Phone =  ifelse(active_dem == 1 & phone != '', 1, 0)) %>%
  group_by(Tier) %>% 
  summarise(Active_Dems = format(round(sum(active_dem)/1000, 0)*1000, big.mark=','),
            High_Propensity = format(round(sum(likely_signer)/1000,0)*1000, big.mark = ","),
            Active_Dem_Phones = format(round(sum(Active_Dem_w_Phone)/1000, 0)*1000, big.mark=','),
            med_age = median(age, na.rm=TRUE), 
            sd_age = sd(age, na.rm=TRUE))
```

### Messaging Campaign

```{r}
## Message testing
df_wide <-left_join(
  df_voters_with_history %>% 
    mutate(age_bracket = cut(age, 
                             breaks = c(18, 35, 50, 65, Inf), 
                             labels = c("18-34", "35-49", "50-64", "65+"),
                             right = FALSE)) 
  , town_agg %>% select(town_name, Tier)) %>%
  filter(party_code %in% c('D') & status == 'A' & phone != '') %>%
  group_by(Tier, age_bracket, gender, likely_signer) %>% 
  summarise(n = n()) %>%
  filter(Tier == 3) %>%
  pivot_wider(
    names_from  = c(gender, likely_signer),
    values_from = n,
    values_fill = 0
  )

tab3 <- df_wide %>%
  ungroup() %>%
  filter(!is.na(age_bracket)) %>%
  mutate(across(contains('_'), ~ format(.x, big.mark = "," ))) %>%
  gt(rowname_col = "age_bracket") %>%
  cols_hide(columns = Tier) %>%

  # Top-level spanners (Gender)
  tab_spanner(
    label = "Male",
    columns = c(M_1, M_0)
  ) %>%
  tab_spanner(
    label = "Female",
    columns = c(F_1, F_0)
  ) %>%
  tab_spanner(
    label = "Unknown",
    columns = c(U_1, U_0)
  ) %>%

  # Second-level labels (Likely)
  cols_label(
    M_1    = "Yes",
    M_0     = "No",
    F_1  = "Yes",
    F_0   = "No",
    U_1 = "Yes",
    U_0 = "No"
  ) %>%

  tab_header(
    title = md("**Tier 3**"),
    subtitle = "Counts by Age, Gender, and Likely Signer Status"
  ) %>%
  # Add border
  cols_align("center") %>%
  opt_all_caps()%>% 
  tab_style(
    style = cell_borders(sides = "right", weight = px(1)),
    locations = cells_body(columns = c(Tier, F_0, M_0)) 
    ) #%>%
  # Add Heatmap
  # data_color(
  #   columns = where(is.numeric),
  #   colors = scales::col_numeric(
  #   palette = c("#f7fbff", "#2171b5"),
  #   domain = NULL
  #   )
  # )


browsable_html <- browsable(
  tagList(
    tags$div(
      style = "display:flex; gap:24px;",
      lapply(
        list(tab1, tab2, tab3),
        function(tbl) HTML(as_raw_html(tbl))
      )
    )
  )
)

save_html(browsable_html, "tiers_side_by_side.html")

```

```{r}
# SCRUBBED PHONES FOR ACTIVE DEMS
scrubbed_phones_active_dems <- read.csv("scrubbed_numbers 2026-01-03 09-24-35.csv")
scrubbed_phones_active_dems$phone_number <- as.character(scrubbed_phones$phone_number)

n_records_with_phone <- as.numeric(table((df_voters_with_history$phone == ''))["FALSE"])
n_records_sharing_phone <- df_voters_with_history %>% group_by(phone) %>% summarise(n=n()) %>% arrange(desc(n)) %>% filter(n > 1) %>% nrow()  
n_unique_phones <- length(unique(df_voters_with_history$phone))
# 309409 records with a phone number 
# 29453 phone numbers assigned to more than one record 
# 274824 unique phone numbers 
# ~10% of phone numbers are shared by more than one record

table(scrubbed_phones$phone_line_type)

vf_active_dems_good_phones <- left_join(df_voters_with_history, scrubbed_phones_active_dems, join_by(phone == phone_number)) %>% filter(!is.na(id.y)) %>%
  filter(fake_number != 'YES') %>%
  filter(phone_line_type == 'CELL PHONE') %>%
  filter(party_code %in% c('D') & status == 'A')

left_join(vf_active_dems_good_phones %>% mutate(age_bracket = cut(age, 
                                                 breaks = c(18, 35, 50, 65, Inf), 
                                                 labels = c("18-34", "35-49", "50-64", "65+"),
                                                 right = FALSE)), 
          town_agg %>% select(town_name, Tier)) %>% group_by(Tier) %>% summarise(n=n())

vf_active_dems_good_phones %>% 
  filter(phone %in% c(vf_active_dems_good_phones %>% group_by(phone) %>% summarise(n=n()) %>% filter(n > 1) %>% pull(phone))) %>%
  arrange(phone) %>%
  select(first_name, last_name, town_name, poll_place_name, phone)
  

w <- left_join(
  vf_active_dems_good_phones %>% 
    mutate(age_bracket = cut(age, 
                             breaks = c(18, 35, 50, 65, Inf), 
                             labels = c("18-34", "35-49", "50-64", "65+"),
                             right = FALSE)) 
  , town_agg %>% select(town_name, Tier)) %>%
  filter(party_code %in% c('D') & status == 'A' & phone != '') %>%
  group_by(Tier, age_bracket, gender, likely_signer) %>% 
  summarise(n = n()) %>%
  pivot_wider(
    names_from  = c(gender, likely_signer),
    values_from = n,
    values_fill = 0
  )

x <- w %>%
  ungroup() %>%
  filter(!is.na(age_bracket)) %>%
  mutate(across(contains('_'), ~ format(.x, big.mark = "," ))) %>%
  gt(rowname_col = "age_bracket") %>%

  # Top-level spanners (Gender)
  tab_spanner(
    label = "Male",
    columns = c(M_1, M_0)
  ) %>%
  tab_spanner(
    label = "Female",
    columns = c(F_1, F_0)
  ) %>%
  tab_spanner(
    label = "Unknown",
    columns = c(U_1, U_0)
  ) %>%

  # Second-level labels (Likely)
  cols_label(
    M_1    = "Yes",
    M_0     = "No",
    F_1  = "Yes",
    F_0   = "No",
    U_1 = "Yes",
    U_0 = "No"
  ) %>%

  tab_header(
    title = md("**Active Dems with Good Phones**"),
    subtitle = "Counts by Age, Tier, Gender, and Likely Signer Status"
  ) %>%
  # Add border
  cols_align("center") %>%
  opt_all_caps()%>% 
  tab_style(
    style = cell_borders(sides = "right", weight = px(1)),
    locations = cells_body(columns = c(Tier, F_0, M_0)) 
    )

save_html(x, "active_dems_with_good_phones_crosstabs.html")

```




### Likely Signers: Demographics and Distributions?

```{r}
# Check Age Distribution
df_voters_with_history %>%
  filter(!is.na(age) & age < 100) %>%
  ggplot(aes(age, fill = as.factor(likely_signer))) +
    #geom_density(position = "identity", alpha = 0.5) +
    geom_histogram(position = "identity", alpha = 0.5, binwidth = 5) +
    scale_fill_brewer(palette = "Set1") +
    theme_minimal() +
    labs(x = "Age", fill = "Likely Signer", y = "Count") +
    ggtitle("Age Distributions: Likely Signers vs. All Other")
    #theme_bw() or theme_grey()

# Check Town Distribution
df_voters_with_history %>%
  group_by(town_name, likely_signer) %>%
  summarise(n = n(), .groups="drop") %>%
  filter(n > 100 & town_name != '') %>%
  group_by(town_name) %>%
  mutate(town_pop = sum(n)) %>% 
  ungroup() %>%
  mutate(perc = paste0(round(100*n/town_pop, 0),'%')) %>%
  ggplot() +
    geom_bar(aes(y = reorder(town_name, n), x = n, fill = as.factor(likely_signer))
             , stat = "identity"
             , position = position_dodge2(reverse=TRUE)) +
    geom_text(data = . %>% filter(likely_signer == 1),   # <- only this category gets labels
              aes(
                y = reorder(town_name, n),
                x = 0,
                label = perc
              ),
              hjust = 1.05,
              size = 4) +
    theme_minimal() +
    labs(x = "Town", fill = "Likely Signer", y = "Count") + 
    ggtitle("Likely Signers by Town")
  
overall_town_pops <- df_voters_with_history %>%
  group_by(town_name) %>%
  summarise(n = n()) %>%
  filter(n> 400) %>%
  ggplot(aes(y=reorder(town_name, n), x=n)) +
  geom_bar(stat = 'identity', position = position_dodge2(reverse=TRUE), fill = 'lightblue') +
  geom_text(aes(y = reorder(town_name, n), x = 0, label = n), color = 'black') +
  ggtitle('Overall Town Populations')
  
  # Looking at proportion of different populations across towns (likely_u35, u_35, and overall_pop)
# likely_signers_u35 <-  df_voters_with_history %>% filter(likely_signer == 1) %>% pull(town_name) %>% table() %>% prop.table() %>% round(2) %>% data.frame() %>% rename(., town = '.', perc_likely_u35 = Freq) 
# left_join(
#   left_join(
#     likely_signers_u35,
#     df_voters_with_history %>% 
#       filter(age <= 35) %>% 
#       pull(town_name) %>% table() %>% 
#       prop.table() %>% round(2) %>%
#       data.frame() %>% 
#       rename(., town = '.', perc_u35 = Freq)
#     , by = 'town'),
#   df_voters_with_history %>% 
#     pull(town_name) %>% 
#     table() %>% prop.table() %>% 
#     round(2) %>% data.frame() %>% 
#     rename(., town = '.', perc = Freq) %>% filter(perc > 0.00),
#   by = 'town'
#   ) %>% arrange(desc(perc_likely_u35))
```








## Key to converting addresses to lat/long

```{r}
# Turn address into lat/lon with tinygeocoder
# Do this with out "likely signer" universe.
df_adds_to_convert <- df_voters_with_history  %>% 
  filter(likely_signer == 1) %>% 
  mutate(full_address = paste(address_number, ' ', street_name, ' ', town_name, ', ', state, ' ', zip5 , sep = ''))
  
batch_size <- 500
batches <- split(df_adds_to_convert, ceiling(seq_len(nrow(df_adds_to_convert)) / batch_size))

results <- map_dfr(batches, ~{
  Sys.sleep(0.2)        # optional
 tidygeocoder::geocode(
      .x,      
      address = full_address,
      method = "census",  # OSM: Uses OpenStreetMap's Nominatim API, Census: Good for well-formatted, U.S. addresses #ArcGIS
      # full_results = TRUE # Gives confidence 0-100
    ) %>%
    rename(lat_census = lat, long_census = long) 
})

# How many of our lat/long coordinates came out of the API as N/A
prop.table(table(is.na(results$lat_census)))

```

When looking at our definition of likely voters, we find that `r nrow(df_adds_to_convert) / df_voters %>% filter(party_code == 'D') %>% nrow()`%
of Democrats in CT-03 meet our definition and `nrow(df_adds_to_convert) / df_voters %>% nrow()`% of all records in the voterfile meet our address. 

```{r}
# Let's try to keep this to points contained within the district
# Option 1: Use the CT-03 Polygon as a border
# Option 2: update the tidygeocoder code to include "full_results = TRUE" to get confidence of the coordinates
# Option 3: identify outliers that are farther than X miles from the centroid
# Option 4: Identify the general box that CT-03 lies in and remove values that are beyond
# Option 5: I found that 99%+ of points are below lat: 41.6, could just use that as a hardcoded filter
# Either way, I would say returning lat/long to NA is preferable to removing the row. 
# Could do this immediately after the batching and then RERUN the batch with OSM or another method to see if those are better for the outliers. 

df_map <- st_transform(df_map, 3857) # Have to make this transformation before joining with the congressional map (needs to be planar or something)
congressional_bound <- st_read("C:/Users/Ruairi/Documents/Coding Projects 2025/ct04_testing/tl_2020_09_cd118")
congressional_bound_003 <- congressional_bound[congressional_bound$CD118FP == '03',] # Grab CT-03 boundaries
congressional_bound_003 <- st_transform(congressional_bound_003, 3857) # Have to make this transformation before joining with town map for consistency

towns_clipped <- st_intersection(df_map, congressional_bound_003) # Makes town map (df_map) respect the congressional boundary

# Remove geographies from the places we clipped out
towns_clipped <- towns_clipped %>% 
  filter(!st_is_empty(geometry))

# Keep only polygon-like geometries
towns_clipped <- towns_clipped %>% 
  st_collection_extract("POLYGON")  # OR "MULTIPOLYGON"

# FOR NOW: Option 5 is the quick fix to get these pushed
##########################################################################

results %>% group_by(lat_census, long_census, full_address) %>% summarise(n=n()) %>% arrange(desc(n))

leaflet_likely_signers <- results %>%
  #filter(town_name %in% towns) %>%
  mutate(under35 = ifelse(age<=35, 1, 0),
         over35 = ifelse(age>35, 1, 0)) %>%
  filter(!is.na(lat_census)) %>% 
  leaflet() %>%
    addTiles() %>%
    addMarkers(
      lng = ~long_census, 
      lat = ~lat_census,
      popup = ~paste0("<b>Address:</b> ", full_address, "<br>",
                "<b>Age:</b> ", age),                  
      clusterOptions = markerClusterOptions(
        # The code below allows us to tweak the color based on the size of the clusters 
        iconCreateFunction = JS("
        function(cluster) {
          var childCount = cluster.getChildCount();

          // Define color thresholds
          var c = 'rgba(0, 175, 30, 0.6)'; // default green
          if (childCount < 200) {
            c = 'rgba(0, 175, 50, 0.6)';      // small
          } else if (childCount < 500) {
            c = 'rgba(255, 180, 0, 0.6)';    // medium = orange
          } else if (childCount < 1000) {
            c = 'rgba(255, 85, 0, 0.6)';    // medium = orange
          } else {
            c = 'rgba(255, 0, 0, 0.7)';      // large = red
          }

          return new L.DivIcon({
            html: '<div style=\"background-color:' + c + '\"><span>' + childCount + '</span></div>',
            className: 'marker-cluster',
          iconSize: new L.Point(40, 40)
          });
        }
      ")
      )
    )

#%>%
#  addCircleMarkers(group = "Group A", ...) %>%
#  addCircleMarkers(group = "Group B", ...) %>%
#  addLayersControl(overlayGroups = c("Group A", "Group B"))  #%>% addProviderTiles(providers$CartoDB.Positron) 

saveWidget(leaflet_likely_signers, "ct03_likely_signers.html", selfcontained = TRUE)

```


## Alt. 1 Mapping points with addGLpoints()

```{r}
# ALT 1: USING addGLPoints
x <- results %>% filter(!is.na(lat_census))
# Have to set the coordinates in order to get the addGLpoints function from leafletextras2 to work 
x_sf <- st_as_sf(x, coords = c("long_census", "lat_census"), crs = 4326)
leaflet() %>%             # add leaflet base map and tiles
  addTiles() %>%
  addGlPoints(
    data = x_sf,              # <-- explicitly pass the filtered dataframe here
    color = "lightblue",
    radius = 5,
    fillOpacity = 0.7
  )
```

## Alt 2. Mapdeck

```{r}
# library(mapdeck)
# set_token("YOUR_MAPBOX_TOKEN")  # get free Mapbox token
# mapdeck(style = mapdeck_style("dark")) %>%
#   add_scatterplot(
#     data = points_df,
#     lat = "lat",
#     lon = "long",
#     radius = 20,
#     fill_colour = "blue",
#     stroke = FALSE
#   )
```

## Alt. 3 Census Tract

```{r}
library(tidycensus)
library(tigris)
options(tigris_use_cache = TRUE)

#census_api_key("YOUR_API_KEY", install = TRUE)
# Convert point data to a shapefile
results_sf <- st_as_sf(results %>% filter(!is.na(lat_census), lat_census < 41.6), coords = c('long_census', 'lat_census'), crs = 4326)

# Grab Census Tract Polygons from Tigris
tracts <- tracts(state = "CT", year = 2023, class = "sf")
tracts <- st_transform(tracts, 4326)

# Join the two, looking for anytime a point falls within a polygon
results_tract_joined <- st_join(results_sf, tracts, join = st_within)

# Aggregate the data by Census Tract to create a count for each tract (drop the point geometry)
census_tract_agg <- results_tract_joined %>% 
                    st_drop_geometry() %>%         # drop point geometry
                    group_by(GEOID) %>%            # group by tract
                    summarise(n_people = n())

# Join the aggregated data back to the Tigris tract data to incorporate the polygon geometry
census_tract_agg_with_geometry <- tracts %>%
  left_join(census_tract_agg, by = "GEOID") %>%
  mutate(n_people = ifelse(is.na(n_people), 0, n_people))
          
# Color palette for choropleth
pal_likely_signers <- colorNumeric("Blues", census_tract_agg_with_geometry$n_people, na.color = "transparent")
          
census_leaflet <- leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(
    data = census_tract_agg_with_geometry %>% filter(n_people > 0),
    fillColor = ~pal_likely_signers(n_people),
    fillOpacity = 0.4,
    color = "black",
    weight = 1,
    popup = ~paste0('Tract: ', NAMELSAD, '<br>',
                    'Likely Signers: ', n_people, '<br>')
  )

saveWidget(census_leaflet, "ct03_likely_signers_census_tract.html", selfcontained = TRUE)
```

